Investigating Determinism of Physics Engines for usage in CAV and V\&V simulations

### 2021-04-13 document/source/body.tex
up to pg.10 TP comments

State-of-the-art game engines provide a convenient option for simulation-based testing. Strong arguments exist that they offer sufficient realism~\cite{Koopman2018} in the physical domain combined with realistic rendering of scenes, potentially suitable for perception stack testing and visual inspection of accidents or near misses. 
%
Furthermore, they are easy to set up and run compared to on-road testing and are simple to control and observe, both with respect to the environment the AV operates in as well as the temporal development of actors~\cite{Ulbrich2015}. 
%
Finally, support for hardware-in-the-loop development or a real-time test-bed for cyber-security testing~\cite{Javaid2013} may also be provided if required. 
%
Compared to the vehicle dynamics simulators and traffic-level simulators used by manufacturers~\cite{FrameworkAndChallenges}, game engines offer a simulation solution that meets many of the requirements for the development and functional safety testing of the autonomous features of AVs in simulation. 
%
However, while game engines are designed primarily for performance to achieve a good user experience, the requirements for AV verification go beyond that and include the issue of determinism.

@@

Investigating the determinism of game engines has not attracted much research interest since performance is more critical for game developers than accurate and repeatable execution. Ensuring software operates deterministically is a non-trivial task.  Catching intermittent failures, or flaky tests~\cite{intermittently-failing-tests}, in a test suite that cannot be replayed makes the debugging process equally difficult~\cite{acm-q-rr-interview}. This section gives an overview of the internal structure of a game engine and what sources or settings in the engine may affect \textit{simulation variance}.

@@

Indeed, the developers of the debugging program \texttt{rr}~\cite{RR_link} took significant steps to ensure deterministic behaviour of their program by executing or context-switching all processes to a single core, which avoids data races as single threads cannot concurrently access shared memory. This allowed control over the scheduling and execution order of threads promoting deterministic behaviour by design~\cite{acm-q-rr-interview}.
%

@@

\begin{table*}[b]
\centering
\begin{tabular}{clclccc}
\toprule
Test & Actors               & Collision    & Collision Type        & $n$  & $\max\sigma$ (m) & $\max\sigma$ (m) \\ 
   &                  &        &               &      & (unrestricted)  & (restricted)  \\ \midrule
1    & Two vehicles                   & No       & N/A               & 1000 & 0.03 & $7.0{\times}10^{-3}$ \\
2    & Two vehicles                   & Yes      & Vehicle and Vehicle     & 1000 & 0.31 & $9.8{\times}10^{-3}$ \\
3    & Two vehicles and a pedestrian  & No       & N/A               & 1000 & 0.07 & $5.2{\times}10^{-4}$ \\
4    & Two vehicles and a pedestrian  & Yes      & Vehicle and Pedestrian    & 1000 & 0.59 & $1.5{\times}10^{-12}$ \\
5    & Two pedestrians                & No       & N/A               & 1000 & $5.6{\times}10^{-13}$ & $5.6{\times}10^{-13}$ \\
6    & Two pedestrians                & Yes      & Pedestrian and Pedestrian & 1000 & $5.6{\times}10^{-13}$ & $5.6{\times}10^{-13}$ \\
\bottomrule
\end{tabular}
\caption{A description of the test scenarios showing the test number, the actors included, whether a collision occurred and if so then between which actors. $n$, the number of repeats is set to 1000 and $\max\sigma$ is the maximum simulation deviation. The term \textit{unrestricted} refers to an unrestricted account of the results including results of any resource utilisation. To understand the impact of collisions and high resource utilisation, the \textit{restricted} column shows a subset of the results where post-collision data and experiments above 75\% resource utilisation have been removed.}
\label{TableOfExperiments}
\end{table*}

@@

@@

\subsection{Process Scheduling Priority} \label{r:process_scheduling}
\begin{figure}[t]
    \centering
    \includegraphics[width=0.48\textwidth]{../other/figures/NICE_analysis_summary_V4.pdf}
    \caption{Variance range of three NICE priority settings for additional CPU \& GPU resource utilisation of 0\% and 75\%.}
    \label{NICEExperimentStressSummary}
\end{figure}

>>>>

State-of-the-art game engines provide a convenient option for simulation-based testing. Strong arguments exist that they offer sufficient realism~\cite{Koopman2018} in the physical domain combined with realistic rendering of scenes, potentially suitable for perception stack testing and visual inspection of accidents or near misses. 
%
Furthermore, they are easy to set up and run compared to on-road testing and are simple to control and observe, both with respect to the environment the AV operates in as well as the temporal development of actors~\cite{Ulbrich2015}. 
%
Finally, support for hardware-in-the-loop development or a real-time test-bed for cyber-security testing~\cite{Javaid2013} may also be provided if required. 
%
Compared to the vehicle dynamics simulators and traffic-level simulators used by manufacturers~\cite{FrameworkAndChallenges}, game engines offer a simulation solution that meets many of the requirements for the development and functional safety testing of the autonomous features of AVs in simulation. 
%
However, while game engines are designed primarily for performance to achieve a good user experience, the requirements for AV verification go beyond that and include \textit{determinism}.

@@

Investigating the determinism of game engines has not attracted much research interest to date since performance is more critical for game developers than accurate and repeatable execution. Ensuring software operates deterministically is a non-trivial task.  Catching intermittent failures, or flaky tests~\cite{intermittently-failing-tests}, in a test suite that cannot be replayed makes the debugging process equally difficult~\cite{acm-q-rr-interview}. This section gives an overview of the internal structure of a game engine and what sources or settings in the engine may affect \textit{simulation variance}.

@@

Indeed, the developers of the debugging program \texttt{rr}~\cite{RR_link} took significant steps to ensure deterministic behaviour of their program by executing or context-switching all processes to a single core, which avoids data races as single threads cannot concurrently access shared memory. This allowed control over the scheduling and execution order of threads, thus promoting deterministic behaviour by design~\cite{acm-q-rr-interview}.
%

@@

\begin{table*}[t]
\centering
\begin{tabular}{clclccc}
\toprule
Test & Actors               & Collision    & Collision Type        & $n$  & $\max\sigma$ (m) & $\max\sigma$ (m) \\ 
   &                  &        &               &      & (unrestricted)  & (restricted)  \\ \midrule
1    & Two vehicles                   & No       & N/A               & 1000 & 0.03 & $7.0{\times}10^{-3}$ \\
2    & Two vehicles                   & Yes      & Vehicle and Vehicle     & 1000 & 0.31 & $9.8{\times}10^{-3}$ \\
3    & Two vehicles and a pedestrian  & No       & N/A               & 1000 & 0.07 & $5.2{\times}10^{-4}$ \\
4    & Two vehicles and a pedestrian  & Yes      & Vehicle and Pedestrian    & 1000 & 0.59 & $1.5{\times}10^{-12}$ \\
5    & Two pedestrians                & No       & N/A               & 1000 & $5.6{\times}10^{-13}$ & $5.6{\times}10^{-13}$ \\
6    & Two pedestrians                & Yes      & Pedestrian and Pedestrian & 1000 & $5.6{\times}10^{-13}$ & $5.6{\times}10^{-13}$ \\
\bottomrule
\end{tabular}
\caption{A description of the test scenarios showing the test number, the actors included, whether a collision occurred and if so then between which actors. $n$, the number of repeats is set to 1000 and $\max\sigma$ is the maximum simulation deviation. The term \textit{unrestricted} refers to an unrestricted account of the results including results of any resource utilisation. To understand the impact of collisions and high resource utilisation, the \textit{restricted} column shows a subset of the results where post-collision data and experiments above 75\% resource utilisation have been removed.}
\label{TableOfExperiments}
\end{table*}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.48\textwidth]{../other/figures/NICE_analysis_summary_V4.pdf}
    \caption{Variance range of three NICE priority settings for additional CPU \& GPU resource utilisation of 0\% and 75\%.}
    \label{NICEExperimentStressSummary}
\end{figure}

@@

\subsection{Process Scheduling Priority} \label{r:process_scheduling}
